---
title: "Problem Set 5"
author: "Sumeet Mishra"
date: "2/26/2019"
output: pdf_document
---
#1.Fit a model that estimates average BMI as a function of height and age. Note that a purely linear model is unlikely to be adequate. Write down the line of code you used to fit your model.

```{r}
library(ggplot2)
library(NHANES)
library(tidyverse)
library(broom)
nhanes<-(NHANES)
myvars1<-c('BMI','Age','Height')
nhanes<-nhanes[myvars1]
nhanes=nhanes[!(!complete.cases(nhanes)), ]
nhanes.lo = loess(BMI ~ Age*Height, data = nhanes, span = 1, normalize=TRUE)
nhanes.plot.df=augment(nhanes.lo)
nhanes.plot.df=nhanes.plot.df%>%filter(Age<=17)
```



#2.Draw a graph faceted by age that shows how, according to your model, average BMI varies with height.

```{r}
ggplot(nhanes.plot.df, aes(x=Height,y=.fitted)) + geom_smooth(method='loess') +
  facet_wrap(~Age ,ncol = 4)+theme_bw()+
  ggtitle('Height~BMI faceted by Age of children less than or equal to 17')+
  xlab('Height in inches')

```

#3.Draw a graph which uses color to denote age that shows how, according to your model, average BMI varies with height. Your color-scheme must be legible and reasonably color- blind friendly.

```{r}
#nhanes1.lo = loess( avg_bmi~ Age+Height, data = nhanes1)
#nhanes1.lo.df = augment(nhanes1.lo)
pallette=c('#D35FB7','#D1E013','#1F9DD4','#006CD1','#E1BE6A','#A1BFAB',
           '#332288','#BB8954','#A7E49B','#E86E6E','#5309F1','#DDCC77',
           '#88CCEE','#44AA99','#117733','#994F00','#E66100')
ggplot(nhanes.plot.df, aes(x =Height , y = .fitted,color=factor(Age)))+
  scale_color_manual(values = pallette) + geom_jitter(height = 0.3, width = 0.1, size = 0.3)+
  geom_smooth(method='loess',method.args=list(degree=1))+
  theme_bw()+labs(color='Age') +
  ggtitle('Height~BMI for children of age less than or equal to 17')+xlab('Height in inches')

```


#4.Draw a graph that uses contours to show how average BMI varies with both height and age.

```{r}
nhanes.grid = expand.grid(Height=seq(80,200,1),Age=seq(2,17,1))
nhanes.predict = predict(nhanes.lo, newdata=nhanes.grid)
nhanes.plot1.df = data.frame(nhanes.grid, fit=as.vector(nhanes.predict))
ggplot(nhanes.plot1.df, aes(x=Height,y=Age,z=fit)) + geom_line() +
  geom_raster(aes(fill = fit))+
  coord_fixed() + scale_fill_distiller(palette="RdYlBu") + 
  geom_contour()+ggtitle('Height~Age for children of age less than or equal to 17')+
  xlab('Height in inches')+ylab('Age in number')
```

#5.How does average BMI vary with height and age for children 17 and under? Is BMI a good measure of body mass for children?
##We oberved from the plots that-
##1.The average BMI and height increases with increase in age. So, young children tends to have less BMI and height than children elder to them.
##2.From the raster and contour plot, we noticed that when age is more than 15, height and BMI tends to be high. 
##3.BMI is not always a good measure of body mass of children as it can not distinguish a obese child from a well built child(child with muscles). There might be other variables for children to be taller or obese which BMI may not explain.












